The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-12 manuscript/manuscript.Rnw
comment out old abstract; clearpage before discussion

The placenta is arguably the most anatomically variable organ in
mammals. Understanding gene expression patterns among placentas from
multiple species can illuminate the molecular basis of placenta
function as well as potentially enrich our knowledge of the biology of
obstetrical syndromes. We profiled the term placenta by using Illumina
RNA-Seq technology to sequence the transcriptomes of
\Sexpr{star.logs[,length(unique(species))]} species of therian (i.e.
marsupials and “placental”) mammals. We identified significant shifts
in gene expression along specific evolutionary lineages using a model
of character evolution. Genes that showed significant elevation in
expression along the human lineage ($\mathrm{FDR}\le 0.01$) were
enriched in GO categories such as angiogenesis, tube morphogenesis,
cell migration, and regulation of inflammatory response. Three of
these genes (KISS1, ADAM12, GPC3) showed exceptionally high expression
in humans; these genes are implicated in preeclampsia and may indicate
that the risk of preeclampsia recently emerged on the human lineage.

@@

>>>>

The placenta is arguably the most anatomically variable organ in
mammals. Understanding gene expression patterns among placentas from
multiple species can illuminate the molecular basis of placenta
function as well as potentially enrich our knowledge of the biology of
obstetrical syndromes. We profiled the term placenta by using Illumina
RNA-Seq technology to sequence the transcriptomes of
\Sexpr{star.logs[,length(unique(species))]} species of therian (i.e.
marsupials and “placental”) mammals.

% We identified significant shifts in gene expression along specific
% evolutionary lineages using a model of character evolution. Genes that
% showed significant elevation in expression along the human lineage
% ($\mathrm{FDR}\le 0.01$) were enriched in GO categories such as
% angiogenesis, tube morphogenesis, cell migration, and regulation of
% inflammatory response. Three of these genes (KISS1, ADAM12, GPC3)
% showed exceptionally high expression in humans; these genes are
% implicated in preeclampsia and may indicate that the risk of
% preeclampsia recently emerged on the human lineage.

@@

\clearpage
